---
title: "model_building"
author: "Wan Ying Hsu"
date: "3/22/2022"
output: word_document
---
#FormCheck - Spam Filtering Development using a Machine Learning Approach

##Data Analysis and Problem Statement
 - Problem: What are the influencing factors that predict web form submission is spam?
 - target variables: label
 - predict variables: 11 continuous variables (flag_count,msg_length, asn_perc, ip_perc, host_perc, domain_perc, name_perc, phone_perc,email_perc,hour, sub_length), 2 categorical variables, 79 binary variables
 
## Data Analysis Proccess
 1. Load data and inspection
 2. Exploratory Data Analysis
 3. Data clearing and reprocessing
 4. Data Partition for training and testing
 5. Features selection
 6. Fit Models and prediction
 7. Model evaluation and comparison

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE,message=FALSE}}
library(cvms) #cross-validation
library(tibble)   
library(DataExplorer) # data visualization
library(dplyr)     # data manipulation
library(tidyverse) # data manipulation
library(rattle) 
library(ROSE) #Random Over-Sampling Examples
library(caret)   # Machine learning
library(randomForest) # Random Forest
library(rpart) # Decision Tree
library(tree)
library(Boruta) #features selection
library(party) #features selection
library(ggplot2) # Data Visualization
library(rpart.plot) # Visualization Decision Tree
library(ROCR) #ROC curve
library(ggpubr)
library(corrplot) 
library(RColorBrewer)
library(glmnet)
```



```{r}
form<-form %>%
  mutate(label = as.factor(label),
         language = as.factor(language),
         server_protocol= as.factor(server_protocol),
         source = as.factor(source))


form <- within(form, {
 language <- 0
 language[language %in% c("cs","de","el","en","es","fr","gs","he","hr","hu","id","it","lv","my","nl","pl","pt","tr")] <- 1
})
form$language <- as.integer(form$language)

form$server_protocol <- recode(form$server_protocol, 'HTTP/1.0' = "HTTP10", 'HTTP/1.1' = "HTTP11", 'HTTP/2.0 '= "HTTP20")
```

```{r}
load('/Users/hsuwanying/Documents/IDS2020_2022/4th_semester/Master_thesis/data/form.rda')
```

```{r}
head(form)
```

```{r}
form<-form %>%
  mutate(label = as.factor(label),
         language = as.factor(language),
         server_protocol= as.factor(server_protocol),
         source = as.factor(source))
```

```{r}
str(form)
```



# Data Exploration
```{r}
introduce(form)
```

```{r}
plot_intro(form)
```
## Graphical analysis
```{r}
ggplot(data = form, aes(x = label, fill = label)) +
  geom_bar() +
  coord_flip() + 
  geom_text(stat = "count", aes(label=..count..),size=3.5,position = position_stack(vjust = 0.5)) 
png("label_distribution.png", width = 15, height = 10, units = "cm", res = 300)
```


```{r}
plot_bar(form, by = 'label')
```

### Continous variables
"flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc",
"name_perc","phone_perc","hour","msg_length", "sub_length"


```{r}
png(file="/Users/hsuwanying/Documents/IDS2020_2022/4th_semester/Master_thesis/qqplot_.png",
width=600, height=350)
plot_qq(form[,c("label","flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")], by = "label")
```


  - use`histogram` to check data distribution
```{r}

plot_histogram(form[,c("label","flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")], nrow = 3)

```
 - `email_perc`, `msg_length`, `sub_length` are right skewed, hence, data scaling is considered

 - use`boxplot` to check outliers
 

```{r}
#png(file="/Users/hsuwanying/Documents/IDS2020_2022/4th_semester/Master_thesis/num_var_histo.png",width=600, height=350)
boxplot(form[,c("label","flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")], by = 'label')
```

```{r}
plot_boxplot(form[,c("label","flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")], by = "label")
```

```{r}
spam <- form[form$label == 'spam',]
ham <- form[form$label == 'ham',]
```


```{r}
plot_histogram(spam, nrow = 3)
```

```{r}
plot_histogram(ham, nrow = 3)
```

```{r}
plot_boxplot(spam, by = 'label')
```

```{r}
plot_boxplot(ham, by = "label")
```


### Using statistic summary to find qutiles
```{r}
summary(form[, c("label","flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")])
```


The histgram, boxplot and statistic summary are used to analyze distribution about continuous variables versue target variable `label`. 
 - the number of flag check `flag_count` in spam spreads out from 3 to 25, whereas, the range is comparatively smaller in ham web forms
 - There are some extreme values found in both messages(msg_length) and subject line(sub_length), the messages length has a maximum 28634 words and 128 in subject line. some outliers also observed when looking at boxplots of name_perc and phone_perc
 

### check predict values `source` vs actual value `label`
```{r}
table(form$source)
table(form$label)
```


## Create Confurion Matrix
```{r}
# make a table of actual value and predict value
tab <- table(Actual = form$label, Prediction = form$source)

conmax <- as_tibble(tab)
conmax
```

```{r}
plot_confusion_matrix(conmax, 
                      prediction_col = "Prediction",
                      target_col = "Actual",
                      counts_col = "n")

```

positive class `label` = 'spam' accounts for 96.8% whereas, `label` = 'ham' only 2.6%, 


### remove predict value column `source`
```{r}
final_df <- subset(form, select = -c(source))
final_df<- final_df %>%
  relocate(label, .before = server_protocol)
head(final_df)
```

# Scale continouse predictors based on output of boxplot, histpgram and summary results
```{r}
num_var <- c("flag_count", "asn_perc", "ip_perc", "domain_perc","email_perc","host_perc","name_perc","phone_perc","hour","msg_length", "sub_length")
```

Logistic regression is one of the Gradient Descent Based Algorithms, it is an 

```{r}
scale_df <- final_df
scale_df[, c(num_var)] = scale(scale_df[, c(num_var)], center = FALSE, scale = TRUE)
plot_boxplot(scale_df, by = 'label')
```

```{r}
boxplot(scale_df[, c(num_var)])
```

```{r}
summary(scale_df[, c(num_var)])
```

```{r}
scale_df <- final_df %>%
  mutate_at(c("email_perc","flag_count", "name_perc", "phone_perc", "msg_length", "sub_length"), ~(scale(.) %>% as.vector))
head(scale_df)  
```

```{r}
plot_boxplot(scale_df, by ='label')
```


### In sample fit all predictor using scale_df
```{r}
logit.all <- glm(label ~., data = scale_df, family=binomial)
summary(logit.all)
```

## Feature Importance using Boruta
Boruta is one of feature selection algorithms, which is capable of working with any classification method that output variable importance measure (VIM). Boruta uses Random Forest. The method performs a top-down search for relevant features by comparing original attributes' importance with importance achievable at random, estimated using their permuted copies, and progressively eliminating irrelevant features to stabilise that test.

References
Miron B. Kursa, Witold R. Rudnicki (2010). Feature Selection with the Boruta Package. Journal of Statistical Software, 36(11), p. 1-13. URL: http://www.jstatsoft.org/v36/i11/
```{r}
citation(package = "Boruta", lib.loc = NULL, auto = NULL)
```

```{r}
boruta <- Boruta(label ~., data = final_df, doTrace =2)
```

```{r}
save(final_df, file = "final_df.RData")
save(boruta,boruta_signif, file = "boruta.RData")
save.image(file = "my_thesis.RData")
```


```{r}
load('/Users/hsuwanying/Documents/IDS2020_2022/4th_semester/Master_thesis/submission/update_report/boruta.RData')
print(boruta)
```


```{r}
boruta_signif <- names(boruta$finalDecision[boruta$finalDecision %in% c("Confirmed", "Tentative")])  
# collect Confirmed and Tentative variables
print(boruta_signif)  # significant variables
```

```{r}
plot(boruta, cex.axis=.5, las=2, xlab="", main="Variable Importance")  # plot variable importance
```

# Get not regeject formular from boruta
74 attributes confirmed important + 8 tentative, a total 82 features are selected
```{r}
formular82 <- getNonRejectedFormula(boruta)
```

```{r}
plotImpHistory(boruta)
```
## Tentative Fix
```{r}
ten.fix <- TentativeRoughFix(boruta)
```

```{r}
print(ten.fix)
```

```{r}
attStats(boruta)
```

# Create Data Partition
```{r}
set.seed(222)
indices <- createDataPartition(scale_df$label, p = 0.8, list = FALSE)
train <- scale_df %>% slice(indices)
test <- scale_df %>% slice(-indices)
```

## check classes of train and test dataset
```{r}
table(train$label)
table(test$label)
```

# Model building

## Logistic Regression Fit all parameters
```{r}
set.seed(1234)
logit.full <- glm(label ~., data = train, family=binomial)
summary(logit.full)
```
###Interptation log.full summary

 - 6 variables: is_server, is_language ,HOSTNAME_NO_REVERSE , MESSAGE_EMPTY , MESSAGE_REPEATING_TEXT,  SENDER_NAME_DIGITS have no information 

```{r}
set.seed(1234)
logit.2 <- update(logit.full, . ~ . -is_server, data = train)
summary(logit.2)
```

```{r}
set.seed(1234)
logit77 <- update(logit78, . ~ . -MESSAGE_EMPTY, data = train)
summary(logit77)
```


# Decision Tree
First fit an unpruned classification tree using all predictors
```{r}
set.seed(1234)
tree.full <- rpart(label ~., data=train, method="class")
tree.full

fancyRpartPlot(tree.full)
rpart.rules(tree.full)
```


```{r}
tree82 <- rpart(formular82, data = train, method="class")

fancyRpartPlot(tree82)
rpart.rules(tree82)
```

```{r}
table(final_df$is_server)
table(final_df$is_language)
table(final_df$HOSTNAME_NO_REVERSE)
table(final_df$MESSAGE_EMPTY)
table(final_df$MESSAGE_REPEATING_TEXT)
table(final_df$SENDER_NAME_DIGITS)
```


# Random Forest
First fit using all predictors
```{r}
set.seed(1234)
rf.full <- randomForest(label~., data = train)
rf.full
```

```{r}
rf84 <- randomForest(label ~ accept_language + msg_length + asn_perc + ip_perc + host_perc + 
    domain_perc + name_perc + phone_perc + email_perc + is_asn + 
    is_hostname + is_domain + is_server + is_language + is_name + 
    is_phone + is_email + is_subject + is_msg + is_weekend + 
    hour + sub_length + msg_long + email_in_msg + url_in_msg + 
    flag_count + MESSAGE_CHAR_METRICS + MESSAGE_NOT_SET + IN_SFS + 
    IP_SYNTAX + HEADERS + MESSAGE_NAME_MATCH + ACCEPT_PATTERN + 
    SENDER_DOMAIN_SERVER_MATCH + ACCEPT_LANG_NULL + HOSTNAME_NO_REVERSE_MATCH + 
    COMPANY_PATTERN + MESSAGE_HOSTNAMES + HOSTNAME_REVERSE_MATCH + 
    HOSTNAME_NO_REVERSE + SUBJECT_RECURRING + NAME_PATTERN + 
    PHONE_PATTERN + MESSAGE_WORD_FREQUENCY + MESSAGE_EMPTY + 
    IP_NO_HOSTNAME + MESSAGE_URL_SPAM + MESSAGE_CYRILLIC_CHARS + 
    IP_IN_SPBL + MESSAGE_URL_PATTERN + SENDER_NAME_PATTERN + 
    MESSAGE_LINKS + MESSAGE_NO_PUNCTUATION + SENDER_EMAIL_REPUTATION + 
    SENDER_EMAIL_SYNTAX + IP_IN_BCC + IP_IN_SPHZ + SENDER_DOMAIN_NO_A + 
    SENDER_DOMAIN_NO_MX + MESSAGE_SHORT + IP_REPUTATION + MESSAGE_CHARACTERS + 
    SUBJECT_CHARACTERS + MESSAGE_REPEATING_WORDS + MESSAGE_REPEATING_TEXT + 
    SUBJECT_PATTERN + SENDER_NAME_SYNTAX + SENDER_HOSTNAME_PATTERN + 
    MESSAGE_PATTERN + SENDER_DOMAIN_SYNTAX + SUBJECT_NON_LATIN + 
    SENDER_FREEMAIL_DOMAIN + MESSAGE_URLS + SENDER_TLD + SENDER_NAME_DIGITS + 
    `HTTP10` + `HTTP11` + `HTTP20` + server_protocol_unknown + language_any + 
    language_de + language_en + language_others + language_unknown, data = train)
rf84
```

```{r}
set.seed(1234)
rf84.100 <- randomForest(label ~ accept_language + msg_length + asn_perc + ip_perc + host_perc + 
    domain_perc + name_perc + phone_perc + email_perc + is_asn + 
    is_hostname + is_domain + is_server + is_language + is_name + 
    is_phone + is_email + is_subject + is_msg + is_weekend + 
    hour + sub_length + msg_long + email_in_msg + url_in_msg + 
    flag_count + MESSAGE_CHAR_METRICS + MESSAGE_NOT_SET + IN_SFS + 
    IP_SYNTAX + HEADERS + MESSAGE_NAME_MATCH + ACCEPT_PATTERN + 
    SENDER_DOMAIN_SERVER_MATCH + ACCEPT_LANG_NULL + HOSTNAME_NO_REVERSE_MATCH + 
    COMPANY_PATTERN + MESSAGE_HOSTNAMES + HOSTNAME_REVERSE_MATCH + 
    HOSTNAME_NO_REVERSE + SUBJECT_RECURRING + NAME_PATTERN + 
    PHONE_PATTERN + MESSAGE_WORD_FREQUENCY + MESSAGE_EMPTY + 
    IP_NO_HOSTNAME + MESSAGE_URL_SPAM + MESSAGE_CYRILLIC_CHARS + 
    IP_IN_SPBL + MESSAGE_URL_PATTERN + SENDER_NAME_PATTERN + 
    MESSAGE_LINKS + MESSAGE_NO_PUNCTUATION + SENDER_EMAIL_REPUTATION + 
    SENDER_EMAIL_SYNTAX + IP_IN_BCC + IP_IN_SPHZ + SENDER_DOMAIN_NO_A + 
    SENDER_DOMAIN_NO_MX + MESSAGE_SHORT + IP_REPUTATION + MESSAGE_CHARACTERS + 
    SUBJECT_CHARACTERS + MESSAGE_REPEATING_WORDS + MESSAGE_REPEATING_TEXT + 
    SUBJECT_PATTERN + SENDER_NAME_SYNTAX + SENDER_HOSTNAME_PATTERN + 
    MESSAGE_PATTERN + SENDER_DOMAIN_SYNTAX + SUBJECT_NON_LATIN + 
    SENDER_FREEMAIL_DOMAIN + MESSAGE_URLS + SENDER_TLD + SENDER_NAME_DIGITS + 
    `HTTP10` + `HTTP11` + `HTTP20` + server_protocol_unknown + language_any + 
    language_de + language_en + language_others + language_unknown, data = train, ntree = 100)
rf84.100
```


# Make predictions
```{r}
# logit.full 
logit.full.pred <-  predict(logit.full, newdata = test,type = "response")
logit.full.Est <- (logit.full.pred > 0.5)
logit.full.True <- (test$label == 'spam')
logit.full.tab <- table(pred = logit.full.Est, truth = logit.full.True)
print()
#logit.full.matrix <- confusionMatrix(logit.full.tab, positive = 'TRUE')
#logit.full.matrix

```

```{r}
tree.full.pred <- predict(tree.full, test, type="class")
tree.full.tab <- table (pred = tree.full.pred, truth = test$label)
tree.full.matrix <- confusionMatrix(tree.full.tab, positive = 'spam')
tree.full.matrix
```


```{r}
rf.full.pred <- predict(rf.full, test)
rf.full.tab <- table(rf.full.pred,test$label)
rf.full.matrix <- confusionMatrix(rf.full.tab, positive = 'spam')
rf.full.matrix
```



```{r}
# logit84
logit84.pred <-  predict(logit84, newdata = test, type = "response")
logit84.Est <- (logit84.pred > 0.5)
logit84.True <- (test$label == 'spam')
logit84.tab <- table(pred = logit84.Est, truth = logit84.True)

logit84.matrix <- confusionMatrix(logit84.tab, positive = 'TRUE')
logit84.matrix

#acccuracy
logit84.acc <- logit84.matrix$overall['Accuracy']
#error rate
logit84.err <-  (1- logit84.acc)
#recall TP/TP+FN => True positive rate
logit84.recall <- 11220/(11220+15)
#precision TP/ TP+FP
logit84.precision <- 11220/(11220+14)
#false positive rate FP/FP+TN
logit84.fpr <- 14/(14+293)
#F1 score 2*(precision*recall/precision +recall)
logit84.f1<- 2*(logit84.precision*logit84.recall/(logit84.precision + logit84.recall))

logit84.df <- data.frame(Accueacy =round((logit84.acc),4), Error_rate = round((logit84.err),4),
                           Recall = round((logit84.recall),4), Precision = round((logit84.precision),4),
                           FP_rate = round((logit84.fpr),4), F1 = round((logit84.f1),4))
logit84.df

```

```{r}
tree84.pred <- predict(tree84, test, type="class")

tree84.tab <- table(pred = tree84.pred, truth = test$label)

tree84.matrix <- confusionMatrix(tree84.tab, positive = 'spam')
tree84.matrix

#Accuracy
tree84.acc <- tree84.matrix$overall['Accuracy']
# error rate
tree84.err <-  (1- tree84.acc)
#recall TP/TP+FN => True positive rate
tree84.recall <- 11222/(11222+24)
#precision TP/ TP+FP
tree84.precision <- 11222/(11222+12)
#false positive rate FP/FP+TN
tree84.fpr <- 12/(12+284)
tree84.f1<- 2*(tree84.precision*tree84.recall/(tree84.precision +tree84.recall))

tree84.df <- data.frame(Accueacy =round((tree84.acc),4), Error_rate = round((tree84.err),4),
                           Recall = round((tree84.recall),4), Precision = round((tree84.precision),4),
                           FP_rate = round((tree84.fpr),4), F1 = round((tree84.f1),4))
tree84.df
```


```{r}
# simple model
logit.simple <- glm(label ~ accept_language + msg_length + asn_perc + ip_perc + host_perc + 
    domain_perc + name_perc + phone_perc + email_perc + is_asn + 
    is_hostname + is_domain + is_server + is_language + is_name + 
    is_phone + is_email + is_subject + is_msg + is_weekend + 
    hour + sub_length + msg_long + email_in_msg + url_in_msg + 
    flag_count + `HTTP10` + `HTTP11` + `HTTP20` + server_protocol_unknown + language_any + 
    language_de + language_en + language_others + language_unknown, data = train, family=binomial)
summary(logit.simple)
```

```{r}
# Build a CART model
CART.simple <- rpart(label ~ accept_language + msg_length + asn_perc + ip_perc + host_perc + 
    domain_perc + name_perc + phone_perc + email_perc + is_asn + 
    is_hostname + is_domain + is_server + is_language + is_name + 
    is_phone + is_email + is_subject + is_msg + is_weekend + 
    hour + sub_length + msg_long + email_in_msg + url_in_msg + 
    flag_count + HTTP10 + HTTP11 + HTTP20 + server_protocol_unknown + language_any + 
    language_de + language_en + language_others + language_unknown, data=train, method="class")
prp(CART.simple)
```

```{r}
set.seed(123)
rf.simple <- randomForest(label ~ accept_language + msg_length + asn_perc + ip_perc + host_perc + 
    domain_perc + name_perc + phone_perc + email_perc + is_asn + 
    is_hostname + is_domain + is_server + is_language + is_name + 
    is_phone + is_email + is_subject + is_msg + is_weekend + 
    hour + sub_length + msg_long + email_in_msg + url_in_msg + 
    flag_count + HTTP10 + HTTP11 + HTTP20 + server_protocol_unknown + language_any + 
    language_de + language_en + language_others + language_unknown, data=train)

rf.simple
```

```{r}
pred.logit.simple <-  predict(logit.simple, type="response")
pred.logit.simple > 0.5

pred.CART.simple <- predict(CART.simple)[,2]
table(test$spam, pred.CART.simple > 0.5)
pred.rf.simple <- predict(rf.simple, type="prob")[,2] 
```











